---
title: "p8105_hw5_qh2284"
author: "Qin Huang"
date: "2023-11-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Problem 1

**Describe the raw data. Create a `city_state` variable (e.g. "Baltimore, MD") and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is "Closed without arrest" or "Open/No arrest").**

```{r}
# load the data
homi = read.csv("./data/homicide-data.csv")
```

The raw data contains `r nrow(homi)` observations and `r ncol(homi)` variables.The dataset contains the id to represent the events, reported_date shows when the events occurred (but this column need to be modified later since it's not readable). It contains the victims' last name, first name, age, race, sex. And it contains where the events happened (city, state, latitude, longitude) and the status of the events. We also could observed that there are some missing values in `lat` and `lon` columns which means we lack of these infos. Also, we found there exists "unknown" in `victim_age`, `victim_sex`, `victim_race`, `victim_last`, `victim_first` to represent missing infos.

```{r}
# create new variable
homi_new = homi %>%
  mutate(city_state = paste0(city,', ',state))

homi_summary = homi_new %>%
  group_by(city_state) %>%
  summarise(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  )

homi_summary
```

**For the city of Baltimore, MD, use the `prop.test` function to estimate the proportion of homicides that are unsolved; save the output of `prop.test` as an R object, apply the `broom::tidy` to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.**

```{r}
library(broom)

bal = homi_summary %>%
  filter(city_state == "Baltimore, MD")

# Calculate the proportion of unsolved homicides
res = prop.test(pull(bal,unsolved_homicides), pull(bal,total_homicides),
          alternative = "two.sided",
          conf.level = 0.95, correct = TRUE) 
res_tidy = broom::tidy(res)

# the result is save in the result folder in the R project
save(res_tidy, file = "result/res_tidy_baltimore.RData")

# pull the estimated proportion and confidence intervals
est_prop = pull(res_tidy, estimate[1])
ci = paste0("(", pull(res_tidy, conf.low),",", pull(res_tidy, conf.high),")")

tidy_frame = data.frame(
  city = 'Baltimore, MD',
  estimate_prop = as.numeric(est_prop),
  ci_low = pull(res_tidy, conf.low),
  ci_high = pull(res_tidy, conf.high),
  confidenct_interval = ci
) 

print(tidy_frame)
```

**Now run `prop.test` for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence intervasl for each. Do this within a "tidy" pipeline, making use of `purrr::map`, `purrr::map2`, list columns and `unnest` as necessary to create a tidy dataframe with estimated proportions and CIs for each city.**

```{r, warning = FALSE}
# run for each city
library(purrr)

# Define the loop_ci function
loop_ci = function(cs) {
  city = homi_summary %>%
    filter(city_state == cs)
  
  res = prop.test(pull(city, unsolved_homicides), pull(city, total_homicides),
                   alternative = "two.sided",
                   conf.level = 0.95, correct = TRUE)
  res_tidy = broom::tidy(res)
  
  est_prop = as.numeric(pull(res_tidy, estimate[1]))
  ci = paste0("(", pull(res_tidy, conf.low), ",", pull(res_tidy, conf.high), ")")
  
  return(data.frame(
                    estimate_prop = est_prop,
                    ci_low = pull(res_tidy, conf.low),
                    ci_high = pull(res_tidy, conf.high),
                    confidence_interval = ci))
}

# Get unique city_state values from homi_summary
cs = unique(pull(homi_summary, city_state))

# Create a dataframe with all unique city_state values
all_cities = data.frame(city_state = cs)

# Use purrr::map and unnest
result_df = all_cities %>%
  mutate(result = map(city_state, loop_ci)) %>%
  unnest(result) 

print(result_df)
```

**Create a plot that shows the estimates and CIs for each city -- check out `geom_errorbar` for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.**

```{r}
result_df %>%
  ggplot(aes(group = city_state, y = reorder(city_state, estimate_prop))) + 
  geom_point(aes(x = estimate_prop)) +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high)) +
  theme_minimal() +
  labs(x = "Estimated proportion with 95% CI", y = "City_state", title = "Estimates and CIs for each city") +
  theme(axis.text.y = element_text(hjust = 0.5,size = 5))

```

From the plot, we could observe the estimated proportion of unsolved with 95% CI for each city, where Chicago, IL has the most and Tulsa, AL has the least.

## Problem 2

**Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time.**

**Start with a dataframe containing all file names; the `list.files` function will help. Iterate over file names and read in data for each subject using `purrr::map` and saving the result as a new variable in the dataframe**

```{r}
all_data = 
  tibble(
    files = list.files("data/problem2/"),
    path = str_c("data/problem2/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()
```

**Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are "tidy", and do any other tidying that's necessary**

```{r}
# tidy the data
tidy_data = 
  all_data %>% mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3),
    ID = str_sub(files, 5, 7)) %>% 
  pivot_longer(week_1:week_8,names_to = "week", values_to = "observation",names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  mutate(observation = as.numeric(observation)) %>%
  dplyr::select(group, ID, week, observation)

tidy_data
```

**Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.**

```{r}
# make a spaghetti plot
spa_plot = tidy_data %>%
  ggplot(aes(x = week, y = observation, group = ID, color = group)) +
  geom_path() +
  geom_point() +
  facet_grid(~group) +
  labs(x = 'Week', y = 'Observation', title = 'Spaghetti plot of observations on each subject over time')

spa_plot
```

Comment: In control group, the observations among weeks for each subject are almost unchanged (fluctuating around a constant). However, in experiment group, the observations for each subject have the trends to increase among weeks.

## Problem 3

**When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected -- put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.**

**First set the following design elements:**

-   **Fix n=30**

-   **Fix** $\sigma = 5$

**Set** $\mu = 0$**. Generate 5000 datasets from the model:** x∼Normal[μ,σ]

```{r}
# run the test
sample = map(1:5000, ~ rnorm(n = 30, mean = 0, sd = 5))
```

**For each dataset, save** $\hat{\mu}$ **and the p-value arising from a test of** H0: $\mu = 0$ **using** $\alpha = 0.05$**. *Hint: to obtain the estimate and p-value, use `broom::tidy` to clean the output of `t.test`.***

```{r}
# define a function to calculate t_test result
t_test = function(sample) {
  result = t.test(sample, alternative = "two.sided", conf.level = 0.95, mu = 0) %>% 
    broom::tidy() %>% 
    dplyr::select(estimate, p_value = p.value)
  return(result)
}

res_list <- map2(sample, 1:5000, ~ t_test(.x) %>%
                   mutate(iteration = .y))

mean_0 = bind_rows(res_list) %>%
  mutate(mean = 0, .before = 1) %>%
  dplyr::select(mean, iteration, everything())

mean_0

```

**Repeat the above for** $\mu = \{1,2,3,4,5,6\}$**, and complete the following:**

```{r}
# create a new function
t_test_new = function(mu) {
  sample = tibble(rnorm(n = 30, mean = mu, sd = 5))
  result = t.test(sample, alternative = "two.sided", conf.level = 0.95) %>% 
    broom::tidy() %>% 
    dplyr::select(estimate, p_value = p.value)
  return(result)
}

mean_1to6 <- expand_grid(mean = 1:6, iteration = 1:5000) %>% 
  mutate(result = map(mean,t_test_new)) %>% 
  unnest(result)
```

```{r}
# combine the data
mean_0to6 = bind_rows(mean_0, mean_1to6)
```

**Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of** μ **on the x axis. Describe the association between effect size and power.**

```{r}
power_plot = mean_0to6 %>%
  group_by(mean) %>% 
  summarize(prop_rejected = sum(p_value < 0.05)/5000) %>% 
  ggplot(aes(x = mean,y = prop_rejected)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(0,6)) +
  labs(x = "True mean",y = "Power",title = "Power vs. true mean") +
  theme_minimal()

power_plot
```

Comment: From above we could observe that as true mean increases, the power also increases. In other words, the power increases as the effect size increases, and ultimately approaches 1. And we could observe that the slope of the curve is decreasing along the true mean.

**Make a plot showing the average estimate of** $\hat{\mu}$ **on the y axis and the true value of** μ **on the x axis.**

```{r}
esti = mean_0to6 %>%
  group_by(mean) %>% 
  summarize(ave_estimate = mean(estimate,na.rm = T), .group = 'drop')

compare_plot = esti %>%
  ggplot(aes(x = mean,y = ave_estimate)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(0,6)) + 
  scale_y_continuous(breaks = seq(0,6)) + 
  labs(x = "True mean",y = "Average estimate mean",title = "Average estimate mean vs. true mean") +
  theme_minimal()

compare_plot
```

Comment: From above, we could observe that the average estimate mean is quite close to true mean. The relationship between average estimate mean and true mean is approximately y = x.

**Make a second plot (or overlay on the first) the average estimate of** $\hat{\mu}$ **only in samples for which the null was rejected on the y axis and the true value of** μ **on the x axis. Is the sample average of** $\hat{\mu}$ **across tests for which the null is rejected approximately equal to the true value of** $\mu$ **? Why or why not?**

```{r}
only_rej = mean_0to6 %>% 
  filter(p_value < 0.05) %>% 
  group_by(mean) %>% 
  summarize(ave_estimate = mean(estimate,na.rm = T), .groups = 'drop')

  
ggplot() +
  geom_line(data = esti, aes(x = mean, y = ave_estimate, color = "esti")) +
  geom_point(data = esti, aes(x = mean, y = ave_estimate, color = "esti")) +
  geom_line(data = only_rej, aes(x = mean, y = ave_estimate, color = "only_rej")) +
  geom_point(data = only_rej, aes(x = mean, y = ave_estimate, color = "only_rej")) +
  scale_x_continuous(breaks = seq(0,6)) +
  scale_y_continuous(breaks = seq(0,6)) +
  labs(x = "True mean",y = "Average estimate mean",title = "Total compared to rejected-only") +
  theme_minimal() +
  scale_color_manual(values = c("esti" = "blue", "only_rej" = "red"))
```

Comment: We could observe that when the true mean is between 1 and 4, the average estimate means of whose null are rejected are different from true mean (always larger). This might be because of the small power at small effect size. When true mean is greater than 4, the average estimate means of whose null are rejected are similar to the true mean.
